---
title: "Individual Project"
author: "Zitan Wang"
date: "2023-09-09"
output: html_document
---

project description:
This project aims to visualize the data related to the COVID-19. current, the plan involves creating:

1) Two global maps accompanied with adjustable timeline to display the monthly cases of death and new cases for each nation.
https://rviews.rstudio.com/2019/10/09/building-interactive-world-maps-in-shiny/#:~:text=Creating%20an%20interactive%20world%20map&text=The%20inputs%20to%20this%20function,the%20aesthetics%20of%20the%20plot.

2) Two scatter plots exploring the degree of different actions taken by each government and the correlated new cases/population proportion in 2021.

```{r setup, warning = FALSE, error = FALSE, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(magrittr)) install.packages("magrittr", repos = "http://cran.us.r-project.org")
if(!require(rvest)) install.packages("rvest", repos = "http://cran.us.r-project.org")
if(!require(readxl)) install.packages("readxl", repos = "http://cran.us.r-project.org")
if(!require(plyr)) install.packages("plyr", repos = "http://cran.us.r-project.org")
if(!require(dplyr)) install.packages("dplyr", repos = "http://cran.us.r-project.org")
if(!require(maps)) install.packages("maps", repos = "http://cran.us.r-project.org")
if(!require(ggplot2)) install.packages("ggplot2", repos = "http://cran.us.r-project.org")
if(!require(reshape2)) install.packages("reshape2", repos = "http://cran.us.r-project.org")
if(!require(shiny)) install.packages("shiny", repos = "http://cran.us.r-project.org")
if(!require(ggiraph)) install.packages("ggiraph", repos = "http://cran.us.r-project.org")
if(!require(RColorBrewer)) install.packages("RColorBrewer", repos = "http://cran.us.r-project.org")
if(!require(countrycode)) install.packages("countrycode", repos = "http://cran.us.r-project.org")
if(!require(wbstats)) install.packages("wbstats", repos = "http://cran.us.r-project.org")
if(!require(tidyr)) install.packages("tidyr", repos = "http://cran.us.r-project.org")
if(!require(lubridate)) install.packages("lubridate", repos = "http://cran.us.r-project.org")
```
##reading in all datasets
```{r}
world_covid_data <- read.csv('./Dataset/full_data.csv')
world_pop_data <- read.csv('./Dataset/master_location_pop_table.csv')
gov_re_data <- read.csv('./Dataset/OxCGRT_latest.csv')
#colnames(gov_re_data)
```


##Import iso3 to match countries
```{r, warning = FALSE, message = FALSE}
library(magrittr)
library(rvest)
library(countrycode)
iso_codes = countrycode::codelist[, c("un.name.en", "iso3c")]
names(iso_codes) = c("Country", "ISO3")
```

##Get the world data with geograpical coordinates directly from package *ggplot2*
```{r, warning = FALSE, message = FALSE}
library(maps)
library(ggplot2)
world_data <- ggplot2::map_data('world')
world_data <- fortify(world_data)

world_data$region[world_data$region == 'USA'] <- 'United States of America'
world_data$region[world_data$region == 'UK'] <- 'United Kingdom of Great Britain and Northern Ireland'
world_data$region[world_data$region == 'North Korea'] <- 'Democratic People’s Republic of Korea'
world_data$region[world_data$region == 'South Korea'] <- 'Republic of Korea'
world_data$region[world_data$region == 'Russia'] <- 'Russian Federation'
world_data$region[world_data$region == 'Laos'] <- 'Lao People’s Democratic Republic'
world_data$region[world_data$region == 'Iran'] <- 'Iran (Islamic Republic of)'

world_data["ISO3"] <- iso_codes$ISO3[match(world_data$region, iso_codes$Country)]
df <- world_data[!duplicated(world_data$region), ]
```


##Clean all three data sets and make new ones, preparing to combine them
```{r}
#processing world_pop_data
world_pop_data_clean <- world_pop_data %>%
 #select(-c(country_alpha_3_code:county_fips_number, data_source_name, geo_location)) %>%
  group_by(country_short_name) %>%
  summarise(geo_latitude = mean(geo_latitude), 
            geo_longitude = mean(geo_longitude), 
            geo_region_population_count = sum(geo_region_population_count)) %>%
  rename('country' = 'country_short_name')
# Only kept 'geo_latitude', 'geo_longitude' and 'geo_region_population_count', all changed to national level. Other data excluded.

####################

#processing world_covid_data
total_covid_death <- world_covid_data %>%
  group_by(location) %>%
  reframe(total_deaths = sum(new_deaths, na.rm = TRUE)) %>%
  rename('country' = 'location')
total_covid_death[total_covid_death == 0] <- NA
# the total number of death for each country

monthly_new_covid <- world_covid_data %>%
  mutate(Period = paste(substr(date, 1, 4), 
                      substr(date, 6, 7), 
                      sep = '')) %>%
  group_by(location, Period) %>%
  reframe('Cases'= sum(new_cases, na.rm = TRUE), 
          'Deaths' = sum(new_deaths, na.rm = TRUE)) %>%
  rename('Country' = 'location')

monthly_new_covid[monthly_new_covid == 0] <- NA
monthly_new_covid["Deaths in thousands"] <- monthly_new_covid$Deaths/1000
monthly_new_covid["Cases in ten thousands"] <- monthly_new_covid$Cases/10000

# Format the date_column to "Mon YYYY"
monthly_new_covid$Period <- ym(monthly_new_covid$Period)
monthly_new_covid$Period <- format(monthly_new_covid$Period, "%b %Y")


# new cases and new death are adjusted to monthly level with 'location' and 'month', other numbers all excluded.
# this data frame still contain data at levels other than national. These data will keep remain here until I match the 'location' in in this data frame with 'country_short_name' in world_pop_data_clean. 'location' not included in 'country_short_name' will be excluded since latitude and longitude are required to plot the global map.

####################

#processing gov_re_data
gov_policies_eval <- gov_re_data %>%
  group_by(CountryName) %>%
  reframe(testing_policy = mean(H2_Testing.policy, na.rm = TRUE), 
          vaccination_policy = mean(H7_Vaccination.policy, na.rm = TRUE)) %>%
  rename('country' = 'CountryName')
gov_policies_eval[gov_policies_eval == 0] <- NA
# I arbitrarily choose only to look at the effects of testing policy and vaccination policy for each country.

####################

#'country' not existed in 'world_pop_data_clean' will be excluded

# In 'total_covid_death', exclude observations that has 'country' not in 'world_pop_data_clean'
total_covid_death <- total_covid_death %>%
  filter(country %in% world_pop_data_clean$country)

# In 'monthly_new_covid', exclude observations that has 'country' not in 'world_pop_data_clean'
monthly_new_covid <- monthly_new_covid %>%
  filter(Country %in% world_data$region)

# In 'gov_policies_eval', exclude observations that has 'country' not in 'world_pop_data_clean'
gov_policies_eval <- gov_policies_eval %>%
  filter(country %in% world_pop_data_clean$country)

# Feed back all the three cleaned data sets to 'world_pop_data_clean' so all four data sets have identical 'country' entries
world_pop_data_clean <- world_pop_data_clean %>%
  filter(country %in% gov_policies_eval$country) %>%
  filter(country %in% total_covid_death$country)
```

##Combine different data sets
```{r}
# create a new data frame named covid_death_with_policy that have total_covid_death and gov_policies_eval combined
covid_death_with_policy <- total_covid_death %>%
  # Adding in world_pop_data_clean
  merge(world_pop_data_clean, by = 'country') %>%
  # Adding in gov_policies_eval
  merge(gov_policies_eval, by = 'country') %>%
  # Death are adjusted to death per million
  mutate(death_per_m = (total_deaths/geo_region_population_count)*1000000) %>%
  filter(not(death_per_m %in% Inf)) %>%
  # Exclude useless columns
  select(-c(total_deaths, geo_region_population_count, geo_latitude, geo_longitude)) %>%
  # Rounding decimals to increase readability
  mutate(across(where(is.numeric), ~round(., 2))) %>%
  mutate(death_per_m = round(.$death_per_m, 0)) 

# %>%
#   rename('Testing Policy' = 'testing_policy') %>%
#   rename('Vaccination Policy' = 'vaccination_policy') %>%
#   rename('Deaths per million' = 'death_per_m') %>%
#   pivot_longer(cols = c('Testing Policy', 'Vaccination Policy'), names_to = 'DataType')

write.csv(covid_death_with_policy, file = "covid_death_with_policy.csv", row.names = FALSE)
####################

# create a new data frame about monthly covid death and cases
monthly_new_covid <- monthly_new_covid %>%
  pivot_longer(cols = c('Cases in ten thousands', 'Deaths in thousands'), names_to = 'DataType') %>%
  rename('Value' = 'value') 

monthly_new_covid['ISO3'] <- iso_codes$ISO3[match(monthly_new_covid$Country, iso_codes$Country)]

# Until now, all data processing have ended. All data frames are cleans. 
# The two of them I will be using to draw plots are 'monthly_covid_data' and 'covid_death_with_policy'
```

## Creating an interactive world map
```{r, warning = FALSE, message = FALSE}
worldMaps <- function(df, world_data, data_type, period){
  
  # Function for setting the aesthetics of the plot
  my_theme <- function () { 
    theme_bw() + theme(axis.title = element_blank(),
                       axis.text = element_blank(),
                       axis.ticks = element_blank(),
                       panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank(),
                       panel.background = element_blank(), 
                       legend.position = "bottom",
                       panel.border = element_blank(), 
                       strip.background = element_rect(fill = 'white', colour = 'white'))
  }
  
  # Select only the data that the user has selected to view
  plotdf <- df[df$DataType == data_type & df$Period == period,]
  plotdf <- plotdf[!is.na(plotdf$ISO3), ]
  
  # Add the data the user wants to see to the geographical world data
  world_data['DataType'] <- rep(data_type, nrow(world_data))
  world_data['Period'] <- rep(period, nrow(world_data))
  world_data['Value'] <- plotdf$Value[match(world_data$ISO3, plotdf$ISO3)]
  
  # Create caption with the data source to show underneath the map
  capt <- paste0("Source: data.world")
  
  # Specify the plot for the world map
  library(RColorBrewer)
  library(ggiraph)
  g <- ggplot() + 
    geom_polygon_interactive(data = subset(world_data, lat >= -60 & lat <= 90), color = 'gray70', size = 0.1,
                                    aes(x = long, y = lat, fill = Value, group = group, 
                                        tooltip = sprintf("%s<br/>%s", ISO3, Value))) + 
    scale_fill_gradientn(colours = brewer.pal(5, "Reds"), na.value = 'white') + 
    labs(fill = data_type, color = data_type, title = NULL, x = NULL, y = NULL, caption = capt) + 
    my_theme()
  
  return(g)
}
```


<!-- ## Creating the two scatter Plots -->
<!-- ```{r, warning = FALSE, message = FALSE} -->
<!-- scatterPlots <- function(df, data_type) -->
<!-- plotData <- df[df$DataType == data_type,] -->
<!-- ``` -->

## Building an R Shiny app
```{r message=FALSE, warning=FALSE}
library(shiny)
library(ggiraph)
shinyApp(
  
  ui = navbarPage("COVID-19 Data",
  tabPanel("World Map",
    fluidPage(
    
    # App title
    titlePanel("Monthly COVID-19 Cases and Deaths Data"),
    
    # Sidebar layout with input and output definitions
    sidebarLayout(
      
      # Sidebar panel for inputs 
      sidebarPanel(
        
        # First input: Type of data
          selectInput(inputId = "data_type",
                    label = "Choose the type of COVID-19 data you want to see:",
                    choices = list("Cases in ten thousands" = "Cases in ten thousands", "Deaths in thousands" = "Deaths in thousands")),
          
          # Second input (choices depend on the choice for the first input)
          uiOutput("secondSelection"),
        
      ),
        
      
      # Main panel for displaying outputs
      mainPanel(
        plotOutput("distPlot")
      )
    )
  )
  ),
  
  ###############################
  
  tabPanel("Testing vs. COVID-19 Deaths",
    fluidPage(
      
      titlePanel("Correlation Between Testing Indicator and COVID-19 Death"),
      
        
      # Main panel for displaying outputs
      mainPanel(
        plotOutput("scatterPlot1")
      )
  )
  ),
  
  ###############################
  tabPanel("Vaccination vs. COVID-19 Deaths",
    fluidPage(
      
      titlePanel("Correlation Between Vaccination Indicator and COVID-19 Death"),
      
      
      # Main panel for displaying outputs
      mainPanel(
        plotOutput("scatterPlot2")
      )
  )
  ),
),
  
  # Define the server
  server = function(input, output) {
    
    # Create the interactive world map
    output$distPlot <- renderGirafe({
      ggplot(code = print(worldMaps(monthly_new_covid, world_data, input$data_type, input$period)))
    })
    
    # Change the choices for the second selection on the basis of the input to the first selection
    output$secondSelection <- renderUI({
      choice_second <- as.list(unique(monthly_new_covid$Period[which(monthly_new_covid$DataType == input$data_type)]))
      selectInput(inputId = "period", choices = choice_second,
                  label = "Choose the period for which you want to see the data:")
    })
    
    output$scatterPlot1 <- renderPlot({
    ggplot(covid_death_with_policy, aes(x = testing_policy, 
                       y = death_per_m)) + 
      geom_point() + 
      theme_minimal() +
      geom_smooth(method = "lm", se = FALSE, color = "red") +
      xlab("Testing Policy") +
      ylab("Deaths per Million") +
      labs(caption = "Source: data.world")
})
    
    output$scatterPlot2 <- renderPlot({
    ggplot(covid_death_with_policy, aes(x = vaccination_policy,
                       y = death_per_m)) + 
      geom_point() + 
      theme_minimal() +
      geom_smooth(method = "lm", se = FALSE, color = "red") +
      xlab("Vaccination Policy") +
      ylab("Deaths per Million") +
      labs(caption = "Source: data.world")
  })
    
  },
   
  options = list(height = 600)
  
)
```

Finally, we can run our app by either clicking "Run App" in the top of our RStudio IDE, or by running
```{r, eval = FALSE}
shinyApp(ui = ui, server = server)
```